
title: "Weather Forecasting with OpenMeteoR"
author: "Taiwo Amodu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weather Forecasting with OpenMeteoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(OpenMeteoR)
```

## Introduction

The `OpenMeteoR` package provides a lightweight, object-oriented interface to access and visualize short-term weather forecasts. It leverages the Open-Meteo API and OpenStreetMap's Nominatim service for geocoding.

Key highlights:
- Retrieve weather forecasts using **coordinates** or **city names**
- Output stored in structured S3 objects: `forecast_result` and `forecast_comparison`
- Custom S3 methods for `print()`, `summary()`, and `plot()`
- Built-in ggplot2 visualizations
- No API key required

This vignette walks you through the main features of the package.

## Installation

If you have a local `.tar.gz` package (e.g., `OpenMeteoR_0.0.0.9000.tar.gz`), install with:

```r
install.packages("OpenMeteoR_0.0.0.9000.tar.gz", repos = NULL, type = "source")
```

Then load the library:

```r
library(OpenMeteoR)
```

---

## 1. Getting Forecast by Location

Use `get_forecast_by_location()` to fetch daily forecasts using city names:

```{r}
result <- get_forecast_by_location("Berlin", days = 3)
```

### Inspect with `print()`

```{r}
print(result)
```

### Summarize with `summary()`

```{r}
summary(result)
```

### Visualize with `plot()`

```{r}
plot(result)
```

The plot shows:
- Max/Min temperatures as lines
- Precipitation as shaded bars

---

## 2. Getting Forecast by Coordinates

Use `get_forecast()` to specify latitude and longitude directly:

```{r}
coords_result <- get_forecast(lat = 48.8566, lon = 2.3522, days = 3)
summary(coords_result)
plot(coords_result)
```

---

## 3. Comparing Multiple Locations

The `compare_forecasts()` function compares forecasts across multiple cities and returns a `forecast_comparison` object.

```{r}
comp <- compare_forecasts(c("Berlin", "Paris", "London"), days = 3)
```

### Inspect

```{r}
print(comp)
```

### Summary

```{r}
summary(comp)
```

### Plot Comparison

```{r}
plot(comp)
```

The comparison plot overlays max temperatures for multiple cities.

---

## 4. S3 Classes Overview

The package defines two custom S3 classes:

### `forecast_result`

- `location`: name of the queried location
- `data`: data frame with columns `date`, `temp_max`, `temp_min`, `precipitation`

Methods:
- `print.forecast_result()`
- `summary.forecast_result()`
- `plot.forecast_result()`

### `forecast_comparison`

- `cities`: vector of city names
- `data`: combined forecast data

Methods:
- `print.forecast_comparison()`
- `summary.forecast_comparison()`
- `plot.forecast_comparison()`

---

## Error Handling

The package handles basic input errors:

- Misspelled locations fall back to closest match from OpenStreetMap
- API connection issues throw informative errors using `rlang::abort()`

---

## Dependencies

- `httr2`: for making API calls
- `jsonlite`: for parsing JSON
- `ggplot2`: for plotting
- `utils`, `stats`: for summary and printing

---

## Conclusion

`OpenMeteoR` demonstrates software-oriented package design using S3 classes and methods to encapsulate and visualize weather forecast data. It is lightweight, user-friendly, and extensible for more advanced analysis or dashboard integration.

---

## Session Info

```{r}
sessionInfo()
```
